# The Monte Carlo Engine {#mc}

```{python simulation}
#| echo: false

import math
import requests
import numpy as np
import matplotlib.pyplot as plt
from datetime import datetime, timedelta, date
import pandas as pd
from scipy.stats import genextreme
from random import choices
import plotly.express as px

NUM_SAMPLES = 10000
SAMPLE_SIZE = 20000

token = pd.read_csv("../tradier.config")
token = token['prod_token'].iloc[0]

def get_hist_data(symbol, 
                  token,
                  start = '2000-01-01', 
                  end = datetime.today().date().isoformat(),
                  endpoint = 'https://api.tradier.com',
                  path = '/v1/markets/history'):

    response = requests.get(f'{endpoint}{path}',
        params = {'symbol':f'{symbol}', 
                  'interval': 'daily', 
                  'start': f'{start}', 
                  'end': f'{datetime.today().date().isoformat()}'},
        headers = {'Authorization': f'Bearer {token}', 
                   'Accept': 'application/json'})
    json_response = response.json()
    
    if type(json_response['history']) != dict:
        return(-1)
    
    hist_price = pd.json_normalize(json_response['history']['day'])
    hist_price = hist_price.rename(({'date': 'Date',
                                     'open': 'Open',
                                     'close': 'Close'}),
                                    axis = 'columns')
    hist_price['perc_change'] = ((hist_price['Close'] - hist_price['Open']) / 
                                    hist_price['Open'])
                                    
    return(hist_price)


def get_current_price(symbol, 
                      token,
                      endpoint = 'https://api.tradier.com',
                      path = '/v1/markets/quotes'):

    response = requests.get(f'{endpoint}{path}',
        params = {'symbols': f'{symbol}', 
                  'greeks': 'false'},
        headers = {'Authorization': f'Bearer {token}', 
                   'Accept': 'application/json'})
    
    json_response = response.json()
    price = json_response['quotes']['quote']['last']
    
    return(price)

def get_latest_day(stock, current_date = datetime.now()):
    if current_date.date().weekday() == 5:
        current_date = current_date.date() - timedelta(days = 1)
    elif current_date.date().weekday() == 6: 
        current_date = current_date.date() - timedelta(days = 2)
    else: 
        if current_date.hour + (current_date.minute / 60) < 9.5:
            current_date = current_date.date() - timedelta(days = 3)
        else:
            current_date = current_date.date()    
    
    return(current_date)


def get_simulation(symbol, 
                   token,
                   option_date,
                   num_samples = NUM_SAMPLES,
                   sample_size = SAMPLE_SIZE, 
                   upper_scale = 0.60, 
                   upper_shape = -0.09, 
                   lower_scale = 0.65, 
                   lower_shape = -0.1,
                   today = datetime.utcnow().date()):

    date = {"Date": pd.date_range(datetime.today().date(), option_date)}

    dates = pd.DataFrame(data = date)
    dates['wday'] = dates['Date'].dt.dayofweek
    dates = dates.loc[dates['wday'] != 5]
    dates = dates.loc[dates['wday'] != 6]

    num_trading_days = dates.shape[0]
    
    dates = dates.reset_index(drop = True)
    
    start = today - timedelta(days = num_trading_days)

    start = start.isoformat()
    start_date = get_latest_day(symbol, datetime.now() - timedelta(days = 1))
    
    hist_price = get_hist_data(symbol, token)
    if type(hist_price) != pd.core.frame.DataFrame:
        return(-1)
    
    perc_pos = 0.5

    perc_neg = 1 - perc_pos

    avg = np.mean(hist_price['perc_change'])
    std = np.std(hist_price['perc_change'])

    samp_upper = genextreme.rvs(upper_shape, 
                                loc = avg, 
                                scale = upper_scale * std,
                                size = round(sample_size * perc_pos))

    samp_lower = -1 * genextreme.rvs(lower_shape, 
                                     loc = avg, 
                                     scale = lower_scale * std,
                                     size = round(sample_size * perc_neg))

    samp = np.append(samp_upper, samp_lower)

    sample_col = pd.Series(range(1, num_trading_days + 1)).repeat(num_samples)
    dates_col = dates['Date'].repeat(num_samples)
    sample_values = np.array(choices(samp + 1, 
                                     k = num_samples * num_trading_days
                                     )).reshape(num_trading_days, 
                                                num_samples)

    start_price = get_current_price(symbol, token)

    sample_values[0] = start_price

    price_paths = sample_values.cumprod(axis = 0)
    price_paths = pd.DataFrame(price_paths)
    price_paths['Date'] = dates['Date']
    
    return(price_paths)


STOCK = 'CMCSA'
options_selection = (date.today() + timedelta(days = 30)).strftime('%Y-%m-%d')
start_date = (date.today() - timedelta(days = 90)).strftime('%Y-%m-%d')
current_price = get_current_price(STOCK, token)
hist_dat = get_hist_data(STOCK, token, start = start_date)
price_paths = get_simulation(STOCK, token, options_selection)

```

```{r create_data}
#| echo: false
#| message: false
#| warning: false


### LIBRARIES
library(reticulate)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)
library(scales)

### DATA CREATION

# DATA SET FOR FULL MC PLOT
many_runs <- py$price_paths %>% 
    pivot_longer(-Date, names_to = "iteration", values_to = "Pred Close") %>% 
    mutate(sim = TRUE, 
           Date = as_date(Date))

mp_plot_dat <- py$hist_dat %>% 
    mutate(sim = FALSE, 
           iteration = as.character(-1),
           Date = as_date(Date)) %>% 
    select(Date, iteration, Close, sim) %>% 
    bind_rows(many_runs)

# DATA SET FOR SINGLE MC PLOT
single_run <- many_runs %>% 
    filter(iteration == 1)

sp_plot_data <- py$hist_dat %>% 
    mutate(sim = FALSE, 
           iteration = as.character(-1),
           Date = as_date(Date)) %>% 
    select(Date, iteration, Close, sim) %>% 
    bind_rows(single_run)

# JUST GET RID OF A GIANT DATASET
many_runs <- NULL

# DATA FOR HISTOGRAM AND BOXPLOT
hist_plot_dat <- mp_plot_dat %>% 
    filter(Date == max(Date))

# GET LAST PRICE 
current_price <- mp_plot_dat %>% 
    filter(sim == FALSE) %>% 
    filter(Date == max(Date)) %>% 
    pull(Close)

# A NUMBER USED IN THE WORDS BELOW AND FICTIONAL STRIKE AND COST
sp <- 42
ep <- 42 + 0.05

count_above <- length(hist_plot_dat$`Pred Close`[hist_plot_dat$`Pred Close` >= ep])
```

```{python delete_big_data}
#| echo: false

# DELETE MORE BIG DATA
del price_paths
```

```{r create_plots}
#| echo: false
#| warning: false

# THE SINGLE RUN PLOT

segment_info <- tibble(hist_y = min(sp_plot_data$Close, na.rm = TRUE) - 0.25,
                       hist_x = sp_plot_data$Date[
                           !is.na(sp_plot_data$Close) & 
                               sp_plot_data$Close == min(sp_plot_data$Close, 
                                                         na.rm = TRUE)],
                       pred_y = max(sp_plot_data$`Pred Close`, na.rm = TRUE) + 0.25,
                       pred_x = sp_plot_data$Date[
                           !is.na(sp_plot_data$`Pred Close`) & 
                               sp_plot_data$`Pred Close` == max(
                                   sp_plot_data$`Pred Close`, na.rm = TRUE)] - 
                           days(1),
                       low_y = min(min(sp_plot_data$`Pred Close`, na.rm = TRUE),
                                   min(sp_plot_data$Close, na.rm = TRUE)) * 0.95,
                       high_y = max(max(sp_plot_data$`Pred Close`, na.rm = TRUE),
                                    max(sp_plot_data$Close, na.rm = TRUE)) * 1.05)

single_run <- ggplot(sp_plot_data, aes(x = Date)) +
    geom_line(aes(y = Close), color = "navy") +
    geom_line(aes(y = `Pred Close`), color = "red") + 
    scale_y_continuous(labels = dollar_format(),
                       limits = c(segment_info$low_y[1],
                                  segment_info$high_y[1])) + 
    ggtitle("Single Run of Monte Carlo Simulation") + 
    annotate("text", 
             label = "Actual\nCMCSA\nPrices", 
             x = segment_info$hist_x[1],
             y = segment_info$hist_y[1],
             vjust = 1, 
             color = "navy") + 
    annotate("text",
             label = "Predicted\nPrice\nChange", 
             x = segment_info$pred_x[1],
             y = segment_info$pred_y[1],
             vjust = 0,
             color = "red") + 
    theme(panel.background = element_blank(),
          panel.grid.major = element_line(color = "light gray"),
          panel.grid.minor = element_blank(),
          axis.ticks = element_blank())

# HERE IS THE PLOT FOR THE FULL MC RUN
segment_info <- tibble(hist_y = min(mp_plot_dat$Close, na.rm = TRUE) - 0.25,
                       hist_x = mp_plot_dat$Date[
                           !is.na(mp_plot_dat$Close) & 
                               mp_plot_dat$Close == min(mp_plot_dat$Close, 
                                                         na.rm = TRUE)],
                       pred_y = mp_plot_dat$Close[
                           mp_plot_dat$Date == max(mp_plot_dat$Date[
                               !is.na(mp_plot_dat$Close)])] * 1.1,
                       pred_x = max(mp_plot_dat$Date[
                               !is.na(mp_plot_dat$Close)]) - days(5))

full_run <- ggplot(mp_plot_dat, aes(x = Date)) +
    geom_line(aes(y = Close), color = "navy") +
    geom_line(aes(y = `Pred Close`, group = iteration), 
              color = "red", alpha = 0.1) + 
    scale_y_continuous(labels = dollar_format()) + 
    ggtitle("Full Monte Carlo Simulation") + 
    annotate("text", 
             label = "Actual\nCMCSA\nPrices", 
             x = segment_info$hist_x[1],
             y = segment_info$hist_y[1],
             vjust = 1, 
             color = "navy") + 
    annotate("text",
             label = "10,000\nSimulated\nPrice\nPaths", 
             x = segment_info$pred_x[1],
             y = segment_info$pred_y[1],
             vjust = 0,
             color = "red") + 
    theme(panel.background = element_blank(),
          panel.grid.major = element_line(color = "light gray"),
          panel.grid.minor = element_blank(),
          axis.ticks = element_blank())

# THIS IS THE HISTOGRAM
histogram <- ggplot(hist_plot_dat, aes(x = `Pred Close`)) +
    geom_histogram(fill = "navy") + 
    geom_vline(xintercept = current_price, color = "red") + 
    geom_vline(xintercept = ep, color = "purple") + 
    xlab("Predicted Prices at Option Expiration") + 
    scale_x_continuous(labels = dollar_format()) +
    annotate("text", 
             label = paste0("Current Price: ", current_price), 
             x = current_price * 0.99, 
             y = 50, 
             hjust = 1, 
             color = "red") + 
    annotate("text",
             label = paste0("Option Effective Price: ", ep), 
             x = ep * 1.01, 
             y = 50, 
             hjust = 0, 
             color = "purple") + 
    theme(panel.background = element_blank(),
          panel.grid.major.y = element_line(color = "light gray"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks = element_blank(),
          axis.title.y = element_blank())

# THIS IS THE HISTOGRAM
ev_ <- mean(hist_plot_dat$`Pred Close`[hist_plot_dat$`Pred Close` >= ep])

ev <- ggplot(hist_plot_dat[hist_plot_dat$`Pred Close` >= ep, ], 
             aes(x = `Pred Close`)) +
    geom_histogram(fill = "navy") + 
    geom_vline(xintercept = ep, color = "purple") + 
    geom_vline(xintercept = ev_, color = "orange") + 
    xlab("Predicted Prices at Option Expiration\nfor Options Above the Effective Price") + 
    scale_x_continuous(labels = dollar_format()) +
    annotate("text",
             label = paste0("Option Effective Price: ", ep), 
             x = ep * 1.001, 
             y = 5, 
             hjust = 0, 
             color = "purple") + 
    annotate("text",
             label = paste0("Expected Price Above Effective Price: ", dollar(ev_)), 
             x = ev_ * 1.001, 
             y = 15, 
             hjust = 0, 
             color = "orange") + 
    theme(panel.background = element_blank(),
          panel.grid.major.y = element_line(color = "light gray"),
          panel.grid.major.x = element_blank(),
          panel.grid.minor = element_blank(),
          axis.ticks = element_blank(),
          axis.title.y = element_blank())

```

I am attempting to predict what the price of any stock will be at an options 
expiration date, some days from now. The Monte Carlo simulation allows me to 
create probabilities for different scenarios. To say that more precisely, the 
simulation allows me to know the probability that an option will be in-the-money 
on its expiration date. 

As an example, I will predict what the price of Comcast (CMCSA) will be 30 days 
from now. 

The simulation is given all of the price change data from CMCSA back to 2000. It 
calculates the daily price changes. The simulation then randomly picks from the 
historical pot of price changes what the stock will do tomorrow and tomorrow and 
tomorrow (and so on, up to the end date 30 days from now). 

```{r print_single}
#| echo: false
#| warning: false
#| fig.align: center
#| dpi: 600

print(single_run)
```

This gets you a single simulated possible path that the stock could take for the 
next 30 days. It is unlikely - probably there is actually 0% chance - that the 
stock will actually follow that exact random path. Here's the magic of the 
simulation: I then create 10,000 to 100,000 more random paths using the exact 
same technique. 

```{r print_full}
#| echo: false
#| warning: false
#| fig.align: center
#| dpi: 600

print(full_run)
```


This full world of simulated price paths allows me to see the distribution of 
potential prices at the option expiration date. I now know the average, median, 
high, low, etc. price at the expiration date. 

More beautifully, I can calculate probabilities that a stock price will reach 
or exceed a strike price at a given options expiration date (more on that below). 

## Effective Price and Being In-The-Money

To find the probability that a certain option is in-the-money, we can simply 
count the number of simulations that finish above (or below for a put) the 
effective price (the strike plus the premium for the option) and divide by the 
total simulations - this is the beauty of the monte carlo simulations.

### Why do we care about the effective price (instead of just Strike Price)? 

It is not enough to find out if an option will only go above its strike price. 
You have to pay to own the option... hence, you are already some money into the 
deal. Since you've already paid some money, you need to make that money back, 
too, to be in-the-money.

When the strike is in-the-money, the premium to buy the option will be large 
enough to make the effective price equal to the current price. The more 
in-the-money, the higher the premium. 

When the strike is out-of-the-money, the premium will be smaller. The further 
out-of-the-money a strike is, the lower the premium. 

### Let's look at an example.

CMCSA is currently trading at `r dollar(current_price)`. Let's use a call option 
with a strike of \$42 that has a premium of \$0.05. The effective price of the 
option (for a call) is: \$42 + \$0.05 = \$42.05. 

What is the probability that we meet or exceed that effective price on the 
strike date? To find that out, we count the number of all of the simulations 
that ended with a price greater than or equal to the effective price and divide 
that by the total count of simulations. 

In this case, we have `r comma(count_above)` simulations that are in-the-money 
out of a total `r comma(py$NUM_SAMPLES)`, or 
`r percent(count_above / py$NUM_SAMPLES, accuracy = 0.01)` chance that this 
option ends in the money.


```{r print_histogram}
#| echo: false
#| warning: false
#| message: false
#| fig.align: center
#| dpi: 600

print(histogram)
```

## Expected Stock Price if In-The-Money

Now that I know the probability of it going above the effective price, I need to 
calculate the expected price of the stock if it is in-the-money. To do this, I 
simply take all simulations that end at or above the effective price and 
calculate the expected value (typically this is the mean, but here I am using 
the median as this dataset DEFINITELY is influenced by outliers).

```{r print_ev}
#| echo: false
#| warning: false
#| message: false
#| fig.align: center
#| dpi: 600

print(ev)
```

## Bringing it all Together: Expected Value of the Option

I have the probability the option will end in-the-money (and by definition the 
probability that the option will end out-of-the-money). I also know the premium 
to buy the option and the expected price of the stock if it is above the 
effective price. I am ready to calculate the expected value of the option. 

```{r final_calc}
#| echo: false

pae <- count_above / py$NUM_SAMPLES
pbe <- 1 - (count_above / py$NUM_SAMPLES)
share_gain <- ev_ - ep
total_gain <- (ev_ - ep) * 100
tgwp <- total_gain - 5

final_worth <- (pae * tgwp) + (pbe * -5)
```

- It costs \$5 to buy the option (100 shares at \$0.05 each)
- There is a `r percent(pae, accuracy = 0.01)` chance 
the option ends in-the-money
- There is a `r percent(pbe, accuracy = 0.01)` chance the option ends 
out-of-the-money
- If the option ends in the money, the stock is expected to be at 
`r dollar(ev_)`, which is equivalent to a gain of `r dollar(ev_)` - 
`r dollar(ep)`, or `r dollar(share_gain)` (times 100 shares)
    - The gain if the option ends in the money is `r dollar(total_gain)` 
    - The final gain, minus the \$5 to buy the option in the first place, is 
    `r dollar(tgwp)`
- Putting all of that together, buying this option is worth 
**`r dollar(final_worth)`**
    - There is a `r percent(pae)` chance you make `r dollar(tgwp)`
    - There is also a `r percent(pbe)` chance you just lose your \$5 premium
    - This expected value is really only possible if you execute this plan 
    many many independent times over the long run
