# Picking Options

Using the simulation outlined on the previous page, I then evaluate options 
from the companies with the highest market capitalization (definition). 

1. Simulate the expected values for all available strike prices (calls and puts) 
for a given options expiration date for all stocks on the list
2. Remove any stock options that do not have a positive expected value
3. Create a few metrics to evaluate each option
4. Pick the "best" option for each stock with a positive expected value
5. Evaluate the cumulative list


### Puts
```{r}
#| echo: false
#| message: false

library(kableExtra)
library(knitr)
library(readr)
library(scales)
library(dplyr)

dis_puts <- read_csv("../dis_puts.csv") 

dis_puts_pres <- dis_puts %>% 
    group_by(symbol) %>% 
    slice_max(EV, n = 1) %>% 
    ungroup() %>% 
    arrange(desc(EV)) %>% 
    mutate(CP = dollar(CP),
           strike = dollar(strike),
           ask = dollar(ask),
           `EP Buy` = dollar(`EP Buy`),
           `Llhd Blw EP` = percent(`Llhd Blw EP`),
           `EV Blw EP` = dollar(`EV Blw EP`),
           EV = dollar(EV)) %>% 
    head(n = 10)

dis_puts_pres %>% 
    kbl() %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
```


### Calls
```{r}
#| echo: false
#| message: false
dis_calls <- read_csv("../dis_calls.csv") 

dis_calls_pres <- dis_calls %>% 
    group_by(symbol) %>% 
    slice_max(EV, n = 1) %>% 
    ungroup() %>% 
    arrange(desc(EV)) %>% 
    mutate(CP = dollar(CP),
           strike = dollar(strike),
           ask = dollar(ask),
           `EP Buy` = dollar(`EP Buy`),
           `Llhd Abv EP` = percent(`Llhd Abv EP`),
           `EV Abv EP` = dollar(`EV Abv EP`),
           EV = dollar(EV)) %>% 
    head(n = 10)

dis_calls_pres %>% 
    kbl() %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
```

I want the most bang for my buck, though. So I created a metric to find out 
which options had the biggest differential between cost and expected value.

```{r}
dis_puts %>% 
    mutate(exp_spent = ifelse(type == "call", 
                              ask * 100 * `Llhd Blw EP`,
                              ask * 100 * (1 - `Llhd Blw EP`)),
           exp_made = ifelse(type == "call", 
                             (`EV Abv EP` - `EP Buy`) * (1 - `Llhd Blw EP`) * -100,
                             (`EV Blw EP` - `EP Buy`) * `Llhd Blw EP` * -100),
           width = exp_made - exp_spent,
           wperc = width / exp_spent) %>% 
    group_by(symbol) %>% 
    slice_max(wperc, n = 1) %>% 
    ungroup() %>% 
    arrange(desc(wperc)) %>% 
    head(n = 10) %>% 
    mutate(CP = dollar(CP),
           strike = dollar(strike),
           ask = dollar(ask),
           `EP Buy` = dollar(`EP Buy`),
           exp_spent = dollar(exp_spent),
           exp_made = dollar(exp_made),
           wperc = percent(wperc)) %>% 
    select(symbol,
           CP,
           type,
           strike,
           ask,
           `EP Buy`,
           exp_spent,
           exp_made,
           wperc)%>% 
    kbl() %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}
dis_calls %>% 
    mutate(exp_spent = ifelse(type == "call", 
                              ask * 100 * (1 - `Llhd Abv EP`),
                              ask * 100 * `Llhd Abv EP`),
           exp_made = ifelse(type == "call", 
                             (`EV Abv EP` - `EP Buy`) *  `Llhd Abv EP` * 100,
                             (`EV Blw EP` - `EP Buy`) * (1 - `Llhd Abv EP`) * 100),
           width = exp_made - exp_spent,
           wperc = width / exp_spent) %>% 
    group_by(symbol) %>% 
    slice_max(wperc, n = 1) %>% 
    ungroup() %>% 
    arrange(desc(wperc)) %>% 
    head(n = 10) %>% 
    mutate(CP = dollar(CP),
           strike = dollar(strike),
           ask = dollar(ask),
           `EP Buy` = dollar(`EP Buy`),
           exp_spent = dollar(exp_spent),
           exp_made = dollar(exp_made),
           wperc = percent(wperc)) %>% 
    select(symbol,
           CP,
           type,
           strike,
           ask,
           `EP Buy`,
           exp_spent,
           exp_made,
           wperc) %>% 
    kbl() %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
```














One thing to consider is the correlation between the stocks. 